---
title: "Homework 7"
author: "Haorui Sun"
date: "3/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Load the dataset
```{r}
library(tidyverse)
data(iris)
glimpse(iris)
```
There are 150 observations and 5 variables: sepal length, sepal width, petal length, petal width, and species.

## 2. Create new data frame `iris1`
```{r}
iris1 <- filter(iris, Species %in% c("virginica", "versicolor"), 
                Sepal.Length > 6, Sepal.Width > 2.5)
glimpse(iris1)
```
There are 56 observations and 5 variables: sepal length, sepal width, petal length, petal width, and species.

## 3. Create new data frame `iris2`
```{r}
iris2 <- select(iris1, Species, Sepal.Length, Sepal.Width)
glimpse(iris2)
```
There are 56 observations and 3 variables: sepal length, sepal width, and species.

## 4. Create new data frame `iris3`
```{r}
iris3 <- arrange(iris2, desc(Sepal.Length))
head(iris3)
```

## 5. Create new data fram `iris4`
```{r}
iris4 <- mutate(iris3, Sepal.Area=Sepal.Length*Sepal.Width)
glimpse(iris4)
```
There are 56 observations and 4 variables: sepal length, sepal width, sepal area, and species.

## 6. Calculation on `iris4`
```{r}
iris5 <- summarize(iris4, avg_sepal_length=mean(Sepal.Length), 
                   avg_sepal_width=mean(Sepal.Width), sample_size=n())
print(iris5)
```

## 7. Calculation on groups from `iris4`
```{r}
iris6 <- summarize(group_by(iris4, Species), 
                   avg_sepal_length=mean(Sepal.Length), 
                   avg_sepal_width=mean(Sepal.Width), sample_size=n())
print(iris6)
```

## 8. Use pipe operation
```{r}
pipe_iris6 <- iris %>%
  filter(Species %in% c("virginica", "versicolor"), 
         Sepal.Length > 6, Sepal.Width > 2.5) %>%
  select(Species, Sepal.Length, Sepal.Width) %>%
  arrange(desc(Sepal.Length)) %>%
  mutate(Sepal.Area=Sepal.Length*Sepal.Width) %>%
  group_by(Species) %>%
  summarize(avg_sepal_length=mean(Sepal.Length), 
            avg_sepal_width=mean(Sepal.Width), sample_size=n())

print(pipe_iris6)
```

## 9. Create the longder data frame
```{r}
longer_iris <- pivot_longer(iris, cols=Sepal.Length:Petal.Width,
                            names_to = "Measure",
                            values_to = "Value")
glimpse(longer_iris)
head(longer_iris)
```
  




