---
title: "Homework6"
author: "Haorui Sun"
date: "2/23/2022"
output: html_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Fit the data distribtuions
### Open libraries
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```

### Read in data vector

The data is about the mass of Steller sea lions *Eumetopias jubatus* at weaning from southeast Alaska.

Data citation:

Hastings, Kelly (2021), Age at weaning in Steller sea lions, Dryad, Dataset, https://doi.org/10.5061/dryad.3bk3j9kh5
```{r}
# extract the column value
df <- read.table("my_dataset.csv",header=TRUE,sep=",")
z <- df$mass
z <- data.frame(1:length(z), z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
summary(z$myVar)
```

### Plot histogram of data
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

### Add empirical density curve
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

### Get maximum likelihood parameters for `normal`
```{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

### Plot `normal` probability density
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", 
                      n = length(z$myVar), args = list(mean = meanML, 
                                                       sd = sdML))
p1 + stat
```

### Plot `exponential` probability density
```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue",
                       n = length(z$myVar), args = list(rate=rateML))
p1 + stat + stat2
```

### Plot `uniform` probability density
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif,
                       colour="darkgreen", n = length(z$myVar), 
                       args = list(min=min(z$myVar), max=max(z$myVar)))
p1 + stat + stat2 + stat3
```

### Plot `gamma` probability density
```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma,
                       colour="brown", n = length(z$myVar),
                       args=list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```

### Plot `beta` probability density
```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) + 
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),
                     start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta,
                             colour="orchid", n = length(z$myVar), 
                             args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

## Simulate the data distributions
The my data best fits the normal distribution, so I would like to simulate a new data set using the normal distribution. 

### Simulate new dataset
```{r}
# find the maximum likelihood parameters
mean <- normPars$estimate["mean"]
sd <- normPars$estimate["sd"]
# simulate the data
len <- length(z$myVar)
newVar <- rnorm(len, mean=mean, sd=sd)
new_z <- data.frame(1:len, newVar)
names(new_z) <- list("ID","myVar")
```

### Plot the new data
```{r}
new_p1 <- ggplot(data=new_z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) +
  geom_density(linetype="dotted",size=0.75)
print(new_p1)
```

### Compare with the original data
```{r}
p1
```
The two histogram profiles look pretty identical, so I think the model is doing a good job of simulating realistic data that match my original measurements. That is because the my realistic data is highly normal distributed.